---
title: "STSCI5080: Group Project -- Chaperters 9, 11, & 12"
author: "Jeff Holden & Will Kretz"
date: "12/1/2017"
output:
  html_document:
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2);library(reshape2)
# setwd("/Class/STSCI 5080/")
```

### Question 1: Dataset Diamonds

Read the data 

```{r diamonds_load}
diamonds2 <- read.table("diamonds2.txt", header=T)
attach(diamonds2)
```

#### 1. Preliminary Data analysis

```{r diamonds_prelim}
summary(diamonds2)
boxplot(X.carat)
boxplot(price)
```

#### 2. (a) Give a 99% confidence interval for the average diamond price.

```{r diamonds_two_a_ci}
t.test(price, conf.level=0.99)
qqnorm(price);qqline(price) 

```
#### 2. (b) Is the average diamond price significantly (at the 1% level) different from 4800?

```{r diamonds_2b_ci }
# Not sure about the results for this
t.test(price, mu=4800, conf.level=0.99)

```

#### 3. Consider the variable carat. Diamonds are considered to be of good quality if the carat weight is above 0.60. Is the average carat weight of the diamonds significantly (at the 5% level) larger than 0.60?

```{r diamonds_three }
t.test(X.carat, mu=0.60, alternative="greater", conf.level=0.95)

```
Yes. The p-value is 0.02563. Therefore, we do not reject the hypothesis that the mean value of the carat weight is abot 0.6.

#### 4. Is the proportion of certificates provided by institute “GIA” significantly (at the 5% level) different from 1/3?

```{r diamonds_four}
gia_cert <- 1*(cert=="GIA")
t.test(gia_cert, mu=1/3, conf.level=1-0.05)
```

Yes. We reject the hypothesis that the true mean is equal to 1/3 at the 5% because the p-value is ~0

#### 5. Are the average weights of the diamonds certified by the three agencies significantly (at the 5% level) different from each other? If so, test, with joint 95% confidence level, which pairs are significantly different.

```{r diamonds_five}
fill <- "#4271AE"
line <- "#1F3552"
pcert <- ggplot(diamonds2, aes(x = cert, y = X.carat)) +
        geom_boxplot(fill = fill, colour = line, alpha = 0.7,
                     outlier.colour = "#1F3552", outlier.shape = 20) +
        scale_y_continuous(name = "Mean carat weight by cert")+
        scale_x_discrete(name = "Cert.") +
        ggtitle("Boxplot of mean carat weight by certification")
pcert <- pcert + theme_bw()
pcert
```

```{r diamonds_five}
 # GIA to HRD
 # GIA to IGI
 # HRD to IGI
 cGIA <- X.carat[cert=="GIA"]
 cIGI <- X.carat[cert=="IGI"]
 cHRD <- X.carat[cert=="HRD"]
 mean(cGIA)
 mean(cIGI)
 mean(cHRD)
 t.test(cGIA,cIGI, alternative="two.sided", conf.level=0.95)
 t.test(cGIA,cHRD, alternative="two.sided", conf.level=0.95)
 t.test(cIGI,cHRD, alternative="two.sided", conf.level=0.95)
```

We reject the null hypohtesis that the means are the same between all three.


### Question 2: Dataset Powernap

```{r powernap_load}
powernap <- read.table("powernap.txt", header=T)
attach(powernap)
```

Was there a difference between 2015 and 2016

```{r powernap_comp}
qqnorm(num2015-num2016); qqline(num2015-num2016)
summary(powernap)
t.test(num2015, num2016, alternative="two.sided", conf.level=0.95, paired=T) # comment on choice paired=T

# ADD A VARIANCE TABLE HERE

```

### Question 3: Dataset set Supermarket

Load the supermarket data
```{r supermarket_load}
supermarket <- read.table("supermarket.txt", header=T)
attach(supermarket)
```

1. Start by performing a preliminary data analysis.


```{r}
summary(supermarket)



super_dt_melt <- melt(supermarket, id = c("display"), variable = c("sales"))
colnames(super_dt_melt) <- c("promotion", "price.reduction", "sales")
super_dt_melt$promotion <- factor(super_dt_melt$promotion)

attach(super_dt_melt)
# JH - FIX X LABELS
super_dt_melt
boxplot(sales~promotion*price.reduction)
```

```{r super_market}
interaction.plot(promotion,price.reduction,sales, main="Intearction for Supermarket Soda Sales")

```
There is very little interaction between the price reduction and promotion. There is a main effect from the promotion and a main effect from the price reduction.

```{r}
promo_aov <- aov(sales~promotion+price.reduction+promotion*price.reduction)
summary(promo_aov)
```
We cannot reject the null, thus there was no significant interaction.

Q3.3

```{r supermarket_three}
promo2_aov <- aov(sales~promotion+price.reduction)
summary(promo2_aov)
str(super_dt_melt)
TukeyHSD(promo2_aov, which="promotion")
plot(TukeyHSD(promo2_aov, which="promotion"))
```

Q3.3

```{r supermarket_three}
promo2_aov <- aov(sales~promotion+price.reduction)
summary(promo2_aov)
str(super_dt_melt)
TukeyHSD(promo2_aov, which="price.reduction")
plot(TukeyHSD(promo2_aov, which="price.reduction"))
```
