---
title: "STSCI5080: Group Project -- Chaperters 9, 11, & 12"
author: "Jeff Holden & Will Kretz"
date: "12/1/2017"
output:
  html_document:
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# setwd("/Class/STSCI 5080/")
```

### Question 1: Dataset Diamondas

Read the data 

```{r diamonds_load}
data <- read.table("diamonds2.txt", header=T)
attach(data)
```

#### 1. Preliminary Data analysis

```{r diamonds_prelim}
summary(data)
boxplot(X.carat)
boxplot(price)
```

#### 2. (a) Give a 99% confidence interval for the average diamond price.

```{r diamonds_two_a_ci}
t.test(price, conf.level=0.99)
qqnorm(price);qqline(price) 

```
#### 2. (b) Is the average diamond price significantly (at the 1% level) different from 4800?

```{r diamonds_2b_ci }
# Not sure about the results for this
t.test(price, mu=4800, conf.level=0.99)

```

#### 3. Consider the variable carat. Diamonds are considered to be of good quality if the carat weight is above 0.60. Is the average carat weight of the diamonds significantly (at the 5% level) larger than 0.60?

```{r diamonds_three }
t.test(X.carat, mu=0.60, alternative="greater", conf.level=0.95)

```
Yes. The p-value is 0.02563. Therefore, we do not reject the hypothesis that the mean value of the carat weight is abot 0.6.

#### 4. Is the proportion of certificates provided by institute “GIA” significantly (at the 5% level) different from 1/3?

```{r diamonds_four}
gia_cert <- 1*(cert=="GIA")
t.test(gia_cert, mu=1/3, conf.level=1-0.05)
```

Yes. We reject the hypothesis that the true mean is equal to 1/3 at the 5% because the p-value is ~0

#### 5. Are the average weights of the diamonds certified by the three agencies significantly (at the 5% level) different from each other? If so, test, with joint 95% confidence level, which pairs are significantly different.

```{r diamonds_five}
# GIA to HRD
# GIA to IGI
# HRD to IGI
# Is this anova?
```

